\documentclass{article}
\usepackage{verbatim,indentfirst,amsmath,amsthm,amssymb,latexsym,zed-csp}
\usepackage[usenames,dvipsnames]{xcolor}

\newcommand{\revise}[1] {\textcolor{red} {#1}} 				% Text to revise


\input{handout}

\begin{document}


\homework{}{August 31, 2016}{Homework \#1: Formal Models and Formal Systems}{}

\noindent\textsc{Note 1}: This homework must be completed using a
LaTex document compiler. So first you will need to install LaTeX and
an editor for it. The course web site has instructions on how to do
this. It also has a template that you should use to create your
answers, as well as examples of LaTex usage
that you might find useful when creating your answers. \vspace {3pt}

\noindent\textsc{Note 2}: For some of these problems (and in general
for the homework in this class), there may be more than one correct
answer.\vspace {3pt}


\begin{enumerate}
% Exercise 1
 \item Consider the game described in Chapter 2 of GWC10. Suppose that
    the container starts out with $N$ balls.
    \begin{enumerate}
        \item List five aspects of the real world that were \emph{not}
        represented in our formal model.
        \item How many ``turns" will it take for the game to stop? Briefly explain why. (Hint: use the $\lfloor$ $\rfloor$ notation to express your solution, if required.)
        \item What is the largest number of extra black balls
        needed, and what configuration of the container causes this number to be required?
        Assume that when two black balls are taken out of the container one is put back into the container and the other into the stock of extra balls.
        \item Argue (informally) that the game stops.
    \end{enumerate}

% Exercise 2
 \item
Consider a language with alphabet $\{ \lambda, \bullet, (, ), x, y, z \}$ and syntax

 \begin{syntax}
     expression  & =  & variable~name | expression, expression\\
     & | & $``$ \lambda $''$, variable~name, $``$ \bullet $''$, expression\\
     & | & ``($"$, expression, ``)$"$;
     \also
     variable~name & = & $``$x$''$ | $``$y$''$ | $``$z$''$;
 \end{syntax}

Are the following wffs of the language? For those that are not briefly explain why.
 \begin{enumerate}
 \item $\lambda x \bullet yz$
 \item $\lambda \bullet x \lambda \bullet y$
 \item $\lambda y \bullet x \bullet z$
 \item $\lambda x \bullet x(yz)$
 \item $\lambda x \bullet \lambda y \bullet xyz$
 \end{enumerate}

% Exercise 3
 \item
Using the {\em Stars} formal system of Example 3.4 from Chapter~3 of GWC10 formally show that\\
 \\
 ${\star}{\diamond}{\star}{\star}{\circ}{\star}{\star}{\star}{\star}{\star}\ {\vdash}\ {\star}{\diamond}{\star}{\star}{\star}{\star}{\circ}{\star}{\star}{\star}{\star}{\star}{\star}{\star}$

% Exercise 4
 \item
 In Chapter~3 of GWC10 the {\em Stars} formal system of Example 3.4 was interpreted as a system for adding certain positive integers. For example, $1 + 3 = 4$ could be proved a theorem of {\em Stars}.

 \begin{enumerate}
 \item Augment {\em Stars} so that you can prove statements such as $3+4=7$ and $15 + 2 = 17$.
 You need to handle only expressions involving the addition of positive integers.

 \textsc{Note:} Your answer should include the alphabet, syntax, inference system, and interpretation.
 \item Show that that $3 + 4 = 7$ is a theorem of the augmented system. (Note: do not forget to provide an interpretation of your result at the end of the derivation process.)
 \end{enumerate}

\end{enumerate}
\end{document}
